---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BLscore

<!-- badges: start -->
<!-- badges: end -->

BLscore provides functions to identify clusters of convergent TCRs 
that are likely to recognize the same epitope. The clustering is based on
pairwise comparisons of all TCRs by aligning their CDR1-3 and calculating 
the alignment scores using BLOSUM62 substitution matrix reflecting evolutionary 
amino acid interchangeability. The three alignment scores (CDR1, CDR2 and CDR3) 
are logistically transformed into a single score (BL-score, BLOSUM-logistic) 
ranging from 0 to 1, where 1 corresponds to the highest probability of specificity 
match and 0 to the lowest. Because the CDRs differ in their impact on TCR
specificity, they are weighted differently in the logistic function. 
The weights and clustering thresholds were established using available data 
sets of TCRs with known specificity (VDJdb and IEDB).

## Installation

You can install the development version of BLscore from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("obrzts/BLscore_private")
```

## Usage

BLscore's main function is `clusterize_TCR` which takes a data.frame with TCR 
sequence data, calculate pairwise BL-scores, uses them to define clusters of
similar TCRs and return a data.frame with cluster ids.

The input data.frame must contain following fields:

* junction_beta - amino acid sequence of CDR3 plus the two flanking conserved 
residues;
* v_beta, j_beta - V and J gene with or without allele; allele information is 
not used for score calculation.

If paired chain clustering is desired junction_alpha, v_alpha and j_alpha must 
be provided too.

An example data.frame is provided with the package:

```{r}
library(BLscore)
head(example_TCR_df)
```

Note, that the default clustering thresholds were defined to optimally detect 
clusters of TCRs recognizing the same epitope. If instead of full junction only 
CDR3 sequence witout flanking residues is provided the scores will be overestimated 
which may lead to wrong cluster assignment.

Usage example:

```{r}
clusters = clusterize_TCR(example_TCR_df, chains="AB", id_col = "id", tmp_folder=".", ncores=4)
head(clusters)
```
